%
% Assignment 3a for CS3530 Computational Theory:
% Regular Languages
% Fall 2017
%
% Problems taken from Sipser
%

\documentclass{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{ae,aecompl}
\usepackage{emp,ifpdf}
\usepackage{graphicx}
\usepackage{enumerate}

\ifpdf\DeclareGraphicsRule{*}{mps}{*}{}\fi

\empaddtoTeX{\usepackage{amsmath}}
\empprelude{input boxes; input theory}

% skip for paragraphs, don't indent
\parskip 6pt plus 1pt
\parindent=0pt
\raggedbottom
\newenvironment{indentlist}{\begin{list}{}{\addtolength{\itemsep}{0.5\baselineskip}}}{\end{list}}

\begin{document}
\begin{empfile}

\begin{center}
\textbf{\Large CS 3530: Assignment 3a} \\[2mm]
Fall 2017
\end{center}

\raggedright

\section*{Problem 1.33 (10 points)}

\subsection*{Problem}

\newcommand{\col}[2]{\genfrac{[}{]}{0pt}{}{\,#1\,}{\,#2\,}}
Let
$$ \Sigma_2=\left\{\col{0}{0},\col{0}{1},\col{1}{0},\col{1}{1}\right\}. $$
Here, $\Sigma_2$ contains all columns of $0$s and $1$s of height
two. A string of symbols in $\Sigma_2$ gives two rows of $0$s and
$1$s. Consider each row to be a binary number and let
$$ C=\{w\in\Sigma_2^*:\text{ the bottom row of }w\text{ is three times the top row}\}$$
For example, $\col{0}{0}\col{0}{1}\col{1}{1}\col{0}{0}\in C$, but
$\col{0}{1}\col{0}{1}\col{1}{1}\notin C$. Show that $C$ is regular.
(You may assume the result claimed in Problem~1.31.)

\subsection*{Solution}

regular languages are closed under reversal and read backwards.lets say 011 = 3, and three times 011 = 1001,  is also equal to the origanl 011 + 0110 = 1001.
Then we can conclude that top row is $T_n$ and bottom for $B_n$, where n = 0,1,2...$n-1$. can be satisfied by this conditions. \newline where $B_n$ = $T_n$ + $T_n-1$.\newline 
we can also determine that $B_n$ values are dependent on $T_n$ values. Then in order to do the DFA to determine if it's regular we have to have a dump state or an unaccepted entire that would not lead to the accept state.
let  a = $\col{0}{0}, b = \col{0}{1}, c= \col{1}{0}, d= \col{1}{1}$

\begin{emp}(0,0)
  t := 2cm;
  node.q0("qa"); q0.c = origin;
  node.q1("qb"); q1.c = q0.c + (t,0);
  node.q2("qd"); q2.c = q0.c + (t,t);
  node.q3("qc"); q3.c = q0.c + (0,t);
  node.q4("q4"); q4.c = q0.c + (2t,.5t);

  makestart(q0);
  makefinal(q0);
  % draw the nodes
  drawboxed(q0,q1,q2,q3,q4);
  
  loop(q0, down,A);
  edge(q0,q1,curve, btex d etex);
  edge(q1,q0,curve,B);
  edge(q1,q2,2,C);
  edge(q1,q4,2,btex a,d etex);
  edge(q0,q4,-65,btex b,c etex);
  edge(q3,q4,80,btex a,d etex);
  edge(q2,q3,curve,A);
  edge(q3,q2,curve,C);
  edge(q2,q4,2,btex bc etex);
  edge(q3,q0,2,B);
  loop(q2,up,D);
\end{emp}


Since the given DFA accepts all strings in $C$ language and rejects all strings that are not in C language, we just proved that $C$ is regular.
% Note: to draw the [0,1] columns in a diagram, use this:
%
%   picture oo, ol, lo, ll;
%   oo = btex $\genfrac{[}{]}{0pt}{}{\,0\,}{\,0\,}$ etex;
%   ol = btex $\genfrac{[}{]}{0pt}{}{\,0\,}{\,1\,}$ etex;
%   lo = btex $\genfrac{[}{]}{0pt}{}{\,1\,}{\,0\,}$ etex;
%   ll = btex $\genfrac{[}{]}{0pt}{}{\,1\,}{\,1\,}$ etex;
%
% Then you can use these as regular labels, e.g.:
%
%   edge(q0,q1,curve,ol);
%
% (0 and 1 are not legal variable names, so I used o and l instead)


\section*{Problem 1.46c (10 points)}

\subsection*{Problem}

Prove that the following languages are not regular. You may use the
pumping lemma and the closure of the class of regular languages
under union, intersection, and complement.

\begin{enumerate}
\item[\bfseries c.] $\{w:w\in\{0,1\}^*$ is not a
palindrome$\}$\footnote{A \textit{\textbf{palindrome}} is a string
that reads the same forward and backward.}
\end{enumerate}

\subsection*{Solution}

Let's assume that $L$ is regular then compliment of the language  is also regular and lets represent that complement language with a $C$, that also means that $C$ is a palindrome language that can be read backwards\newline
Let $S$ a string = xyz $\in$ $L$. $S$ = 001 \newline
Let x= $0^p$, y= $0^k$, z= $1^p$. and $k$,$p$ $>$ 0.\newline
Then we can conclude that $S$ is the not the same string backwards since $k-p < p$. and it does not belong to $C$ language. by the pumping lemma we can say that $C$ is not a regular language. and therefore L is not a regular language.

\end{empfile}
\immediate\write18{mpost -tex=latex \jobname}
\end{document}
